Towards Comparable Active Learning

### 2023-02-21 doc/al_benchmark_paper.tex
upper bound

\subsection{Contribution}
\begin{itemize}
	\item Benchmark suit for Active Learning
	\item Properly optimized classification models
	\item Multiple use-cases: Single Dataset, Dataset Transfer
	\item Multiple Domains: Tabular, Image, Text
\end{itemize}

@@

\subsection{Problem Description}
We assume a dataset $\mathcal{D} := (\mathcal{U}, \mathcal{L})$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set with $|\mathcal{L}| \ll |\mathcal{U|}$. 
For the purpose of this work, only fully labeled datasets are used and the labels for $\mathcal{U}$ are suppressed until they are selected for labeling.
For evaluation purposes we also assume a held-out test set $(x_{test}, y_{test}) \in (\mathbb{R}^M, \mathbb{R}^C)$
We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{L}$.
To perform classification, a model $\phi_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used. To fit the model, it is parameterized by $\theta$ and subjected to loss $l(y, \phi_\theta(x)) := \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$. For this work, categorical cross-entropy (CE) is used.
For evaluating classification performance, we use Accuracy \\ [1mm]
Active learning is defined as sequentially selecting single instances of $x_i \in \mathcal{U}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L} := \mathcal{L} \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ normalized AUC:
\begin{equation}
	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \phi, B) := \frac{1}{B} \sum_{i=1}^{B} acc(y_{test}, \phi_i(x_{test}))
\end{equation}
Where $\phi_i$ is the retrained classification model after the i-\textit{th} instance was selected

>>>>

\subsection{Contribution}

@@

\subsection{Problem Description / Delineation}
\textbf{Basic Classification}\\
We assume a dataset $\mathcal{D} := (\mathcal{U}, \mathcal{L})$ consisting of a large pool of unlabeled data $\mathcal{U}$ and a small set of labeled data called the seed set with $|\mathcal{L}| \ll |\mathcal{U|}$. 
For the purpose of this work, only fully labeled datasets are used and the labels for $\mathcal{U}$ are suppressed until they are selected for labeling.
For evaluation purposes we also assume a held-out test set $(x_{test}, y_{test}) \in (\mathbb{R}^M, \mathbb{R}^C)$
We consider only classification problems, hence the instances of a dataset have the form $x \in \mathbb{R}^M$ for $\mathcal{U}$ and $(x, y) \in (\mathbb{R}^M, \mathbb{R}^C)$ for $\mathcal{L}$.
To perform classification, a model $\hat y_\theta := \mathbb{R}^M \rightarrow \mathbb{R}^C$ is used. To fit the model, it is parameterized by $\theta$ and subjected to loss $l(y, \hat y_\theta(x)) := \mathbb{R}^C \times \mathbb{R}^C \rightarrow \mathbb{R}$. For this work, categorical cross-entropy (CE) is used.
For evaluating classification performance, we use Accuracy \\ [1mm]
%
\textbf{Pool-based AL with single instances (non-batch setting)}\\
Active learning is defined as sequentially removing single instances $x_i \in \mathcal{U}_i\hspace{1mm}; \hspace{2mm} \mathcal{U}_{i+1} := \mathcal{U}_i / \{x_i\}$, requesting their label $y_i$ and adding them to the labeled pool $\mathcal{L}_{i+1} := \mathcal{L}_i \cup (x_i, y_i)$ until a fixed budget $B$ is exhausted $i := 0 \ldots B$.
After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on a held-out test set.
The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}
	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} acc(y_{test}, \hat y_i(x_{test}))
\end{equation}
Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected. \\ [1mm]
%
\textbf{Framing AL as RL}\\
We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $x_i \underset{\tau}{\sim} \mathcal{U}_i$, constructs the state $s_i := \omega(x_i)$ and presents it to the agent to select an action $a_i := \Omega(s_i)$.
The action $a_i$ is the index of the selected instance in $u_i$ out of all possible indices $A := [0 \ldots \tau]$.
This process is repeated $B$ times $i := [0 \ldots B]$.
